A Quantitative Genetic Linked Selection Model of the Negative Selection Process in Humans

### 2023-07-20 manuscript/manuscript.tex
ms updates

XXX HERE

@@

@@

>>>>

@@

\section*{Locally-rescaled Fits}

Given that our simulations indicated that $B(x)$ is more accurately predicted
when we use local-rescaling, we modified our composite-likelihood method to be
run a second time. Intuitively, this is based on the notion that if a neutral
allele experiences an effective population size of $B(x)N$, so too should a
selected allele, and this should be considered in how the SC16 equations are
solved. Immediately, there is an important caveats to note about this analysis.
Local-rescaling is done using the first round of maximum likelihood estimates,
and then re-solving the system of equations with a population size of $B(x)N$.
This could introduce the possibility of some circularity, since the $B(x)$ at
this position includes the contribution of the focal segment that is being
rescaled. However, the qualitative results do not change based on whether the
$B(x)$ is at averaged at the 1 kbp level or at the megabase scale; for the
latter, a single segment makes little contribution to the average.

With this caveat noted, there are three important but tentative results to draw
from this analysis. First, the estimated mutation rate is implausibly higher,
reaching the upper limit of our optimization boundary of $\mu = 8 \times
10^{-8}$ across all populations in or PhastCons Priority model (Supplementary
Material XXX). Second, all of our leave-one-chromosome-out $R_\text{LOCO}^2$
are about one percentage point higher than the unrescaled model. This improved
fit could be an artifact of the circularity, so it is unimportant. Third, the
DFE estimates for PhastCons regions is now U-shaped, with $75-77\%$ of mass
being placed on a weakly deleterious class, $s=10{^-5}$. Interestingly, this is
the first of all of our models where such an appreciable mass has been placed
on a midpoint in our grid; in all other cases, non-strongly estimates were
neutral ($s=10^{-8}$). Finally, predicted pairwise diversity is nearly
identical to our original, non-locally-rescaled fit (see Supplementary
Materials XXX). This

Overall, this leads to an open end in our work. If selected sites along the
genome were to experience a draft effective population size that is closer to
our locally-rescaled version, then these quantitative genetic models of linked
selection

the fit to the linked selection signal is very close under
different parameter combinations (Supplementary Materials XXX). The fits with
local-rescaling

Intuitively, this is based on the notion that if a neutral allele experiences
an effective population size of $B(x)N$, so too should a selected allele. We
find in simulations this reduces error around $2 N_e s \approx 1$.
Unfortunately, employing this in our composite-likelihood method would create
some circularity, since the local $B(x)$ experienced at a conserved segment
includes the reduction due to the segment itself. There are not computationally
feasible ways to avoid this issue in the current implementation.

@@

Linked selection processes alter genetic diversity on multiple scales, creating
patterns based on their strength relative to the stochasticity of coalescence
times and mutations. Sweeps typically leave signals on shorter genomic scales
(i.e. kilobases; \cite{Akey2004-zg}) which combined with the paucity of
adaptive substitutions in humans suggests positive selection is generating the
apparent large, megabase-scale linked selection signal
\parencite{Hernandez2011-gs,Murphy2022-sj}. By contrast, there is reasonable
evidence that negative selection against new mutations occurring in
highly-conserved sequences is generating large-scale chromosomal patterns of
diversity.

Population genetic theory has been productively applied to fit this linked
selection signal, producing plausible estimates of the strength of selection at
these conserved regions and mutations rates consistent with more precise
pedigree-based estimates. However, these models rely on the canonical
background selection model, which due to its assumption that mutations are
sufficiently deleterious they cannot fix, has constrained statistical methods
to only fit background selection models in regions \emph{a priori} under
constraint. In reality, the relationship between diversity levels and the
strength of selection at deleterious mutations is U-shaped, which implies there
could be more uncertainty that previously appreciated in the distribution of
weak and strongly deleterious mutations.

We sought to address this uncertainty in this work, by leveraging a different
class of linked selection models based on modeling the equilibrium fitness
variance \parencite{Santiago2016-mu}. This model is more general than classic
background selection theory; the background selection model is analytically a
special case when the deleterious substitution rate is zero. Moreover, the SC16
model explains equilibrium fitness variance across all selection coefficients
by jointly predicting another central quantity in evolutionary genomics: the
deleterious substitution rate in conserved features. This offers a further way
to confirm or falsify that our understanding of the negative selection process
is accordance with other observed quantities.

However, while our improved negative selection model leads to nearly identical
parameter estimates as previous working employing classic BGS theory, we find a
fundamental contradiction in the implied rate of deleterious substitution and
observed divergence in conserved regions. Unfortunately, this suggests a
potentially serious model misspecfication or non-identifiability issue. Since
the rate of substitutions is determined by the fraction of DFE mass below $2N_e
s \le 1$, the consistency between our DFE estimates and those from
\textcite{Murphy2022-sj} suggests this discrepancy is shared across models of
negative selection.

One clue towards solving this contradiction is given by our simulation work.
While our linked selection model accurately predicts the substitution rates,
fitness variance, and diversity for $2N_e s > 1$ and $2N_e s \ll 1$, there's a
nominal error around the drift boundary of $2N_e s \approx 1$. This region of
parameter space is at a delicate balance between stochasticity dominating
allele frequency dynamics (for $2N_e s \ll 1$) and having little influence
($2N_e s > 1$). Our implementation of the SC16 model in essence only considers
the reduction due to selection \emph{within} the segment and not neighboring
regions. Thus, like classic BGS methods, our model does not consider selective
interference due to neighboring regions under selective constraint. Solving the
SC16 equations using a locally-rescaled population size of $B(x) N$ rather than
$N$ are in effect approximating the reduction in effective population due to
selective interference. Intuitively, this is based on the notion that if a
neutral allele experiences an effective population size of $B(x)N$, so too
should a selected allele. We find in simulations this reduces error around $2
N_e s \approx 1$. Unfortunately, employing this in our composite-likelihood
method would create some circularity, since the local $B(x)$ experienced at a
conserved segment includes the reduction due to the segment itself. There are
not computationally feasible ways to avoid this issue in the current
implementation.

Out of curiosity, we experimented with fitting human diversity data, using the
predicted $\widehat{B}(x)$ for local-rescaling, and rerunning our
composite-likelihood framework (see Supplementary Mathematical XXX). While the
circularity issue means these results should be interpreted with caution, we
generally find these models fit nearly as well

it creates a bit of circularity.
